
```{r}
# Replace with your path

path_root<-"G:/Meine Ablage/AAA_MasterThesis12_08/turn_in/"

```


I created this script as a notebook, so each chunk could be run individually. Therefore, the "setwd()" command does not work for all chunks.  




### 1. Load Libraries
```{r message=FALSE, warning=FALSE}
library(scales)
library(gsignal)
library(bfast)
library(caret)
library(ggh4x)
library(colorspace)
library(dplyr)
library(feasts)
library(forecast)
library(fpp3)
library(ggplot2)
library(ggpubr)
library(imputeTS)
library(isotree)
library(keras)
library(lubridate)
library(leaflet)
library(plotly)
library(purrr)
library(raster)
library(sf)
library(sp)
library(tidyverse)
library(terra)
library(tensorflow)
library(tsibble)
library(zoo)
filter<-dplyr::filter
select<-dplyr::select
mutate<-dplyr::mutate
```


### 2.Loading Test and Training Data 
```{r}

path1<-paste0(path_root, "Data/prepared_data/Final_Data/prepared data/conifer/train/sg_ma_train.csv")
sg_ma_interpolated <- read.csv(path1)

path2<-paste0(path_root, "Data/prepared_data/Final_Data/prepared data/conifer/train/ma_train.csv")
ma_interpolated <- read.csv(path2)

path3<-paste0(path_root, "Data/prepared_data/Final_Data/prepared data/conifer/train/kalman_train.csv")
kal_interpolated <- read.csv(path3)

path4<-paste0(path_root,"Data/prepared_data/Final_Data/prepared data/conifer/health-soph-id6-combined_PIP_decid_removed.csv")
test_data <- read.csv(path4)

# Adding index changes that have been applied in the preprocessing script but not to this data
ma_interpolated <- ma_interpolated |> indices()
sg_ma_interpolated <- sg_ma_interpolated |> indices()
kal_interpolated <- kal_interpolated |> indices()
test_data <- test_data |> indices()

### 3. Definition of the applied functions
```

### 3. Definition of the Applied Functions
```{r}
#functions


long2tens <- function(x) {
  x <- predict(scaler, x)  # Scaling the dataframe 
  x <- as.matrix(x)
  ts <- array_reshape(x, c(dim(x)[1], 1, dim(x)[2]))
  ts <- tf$convert_to_tensor(ts, dtype = tf$float32)
}

long2tens3d_no_scaler <- function(x, n_samples) {
  x <- as.matrix(x)
  timesteps_per_sample <- nrow(x) / n_samples
  
  if (nrow(x) %% n_samples != 0) {
    stop("The number of samples does not divide the time steps evenly.")
  }
  
  ts <- array_reshape(x, c(n_samples, timesteps_per_sample, ncol(x)))
  ts <- tf$convert_to_tensor(ts, dtype = tf$float32)
  
  return(ts)
}

array2df <- function(x) {
  x <- array_reshape(x, c(dim(x)[1], dim(x)[3]))
  x <- as.matrix(x)
  x <- as.data.frame(x)
}

fill_gaps_ma <- function(ts) {
  ts_clean <- ts |> 
   # dplyr::select(all_of(names)) |> 
    mutate(time = yearweek(time)) |> 
    group_by(time) |> 
    summarise_all(mean) |> 
    as_tsibble() |> 
    fill_gaps() |> 
    na_seadec(algorithm = "ma", find_frequency = TRUE) |> 
    as.data.frame() |> 
    mutate(time = as_date(time))
  
  return(ts_clean)
}

fill_gaps_kal <- function(ts) {
  ts_clean <- ts |> 
   # dplyr::select(all_of(names)) |> 
    mutate(time = yearweek(time)) |> 
    group_by(time) |> 
    summarise_all(mean) |> 
    as_tsibble() |> 
    fill_gaps() |> 
    na_seadec(algorithm = "kalman", find_frequency = TRUE) |> 
    as.data.frame() |> 
    mutate(time = as_date(time))
  
  return(ts_clean)
}

fill_gaps_sg_ma <- function(x) {
  x <- x |> 
    as.data.frame() |> 
    mutate(time = yearweek(time)) |> 
    group_by(time) |> 
    summarise_all(mean, na.rm = TRUE)
  t <- x |> dplyr::select(time)
  x <- x |> 
    dplyr::select(-time) |> 
    as.matrix() |> 
    sgolayfilt() |> 
    cbind(t, x) |> 
    as_tsibble(index = time) |> 
    fill_gaps() |> 
    as.data.frame() |> 
    na_seadec(algorithm = "ma") |> 
    mutate(time = as_date(time)) |> 
    as.data.frame()
  
  return(x)
}  
```



### 4. Defining the Parameters of the Model

The following parameters are defined in the upcomming code chunk:

+ **variables to use**
+ **length of the input sequences** 
+ **imputation method** 


The model architectures are: 
- **Model 1:** Three LSTM layers with 128, 64, and 8 units.
- **Model 2:** Three LSTM layers with 256, 64, and 8 units.
- **Model 3:** Two LSTM layers with 16 and 4 units.
- **Model 4:** Three LSTM layers with 128, 32, and 8 units.
- **Model 5:** Two LSTM layers with 128 and 8 units.


```{r}
# Custom Attention-Layer: zeitliche Gewichte (softmax) und Summen-Pooling über Zeit
layer_attention_sum <- new_layer_class(
  classname = "AttentionSum",
  
  initialize = function(...) {
    super()$`__init__`(...)
  },
  
  build = function(input_shape) {
    # input_shape: list(batch, timesteps, features)
    timesteps <- as.integer(input_shape[[2]])
    features  <- as.integer(input_shape[[3]])
    
    self$W <- self$add_weight(
      name = "att_weight",
      shape = c(features, 1L),
      initializer = initializer_random_normal(stddev = 0.05)
    )
    self$b <- self$add_weight(
      name = "att_bias",
      shape = c(timesteps, 1L),
      initializer = initializer_zeros()
    )
    super$build(input_shape)
  },
  
  call = function(x, mask = NULL, training = NULL) {
    K <- backend()
    # et = tanh(x·W + b), dann softmax über die Zeitachse
    # x: [B, T, F]; W: [F,1] -> x·W: [B,T,1]
    et <- K$tanh( K$dot(x, self$W) + self$b )      # [B, T, 1]
    et <- K$squeeze(et, axis = -1L)                # [B, T]
    
    # Mask optional beachten: mask ist [B, T] (TRUE/FALSE)
    if (!is.null(mask)) {
      # setze unmaskierte Positionen extrem negativ vor softmax
      minus_inf <- K$constant(-1e9)
      et <- K$switch(
        K$cast(mask, dtype = K$dtype(et)),
        et,
        minus_inf
      )
    }
    
    at <- K$softmax(et)                            # [B, T]
    at <- K$expand_dims(at, axis = -1L)            # [B, T, 1]
    
    output <- x * at                               # [B, T, F]
    K$sum(output, axis = 2L)                       # Summe über Zeit -> [B, F]
  },
  
  compute_output_shape = function(input_shape) {
    # -> [batch, features]
    list(input_shape[[1]], input_shape[[3]])
  },
  
  get_config = function() {
    c(super$get_config())
  }
)

# Bequemer Layer-Wrapper für das Pipe-API
layer_attention_sum <- function(object, name = NULL) {
  create_layer(layer_attention_sum, object, list(name = name))
}




```





```{r}

# Defining variables
vars1 <- c("time", "nir", "red", "rededge1", "rededge2", "rededge3", "blue", "green", "swir16", "swir22")
vars2 <- c("time", "ndvi", "s2rep", "tcw", "dwsi")
vars3 <- c("time", "rededge1", "rededge2", "rededge3", "nir")
vars4 <- c("time", "ndmi", "evi2", "s2tcw", "s2tcg", "dwsi", "s2rep", "srswir")
vars5 <- c("time", "s2tcw", "s2tcg", "s2tcb")

# Model definitions

attention_model <- function(time_steps, n_variables,
                            bottleneck = 64,
                            out_activation = c("sigmoid","linear","tanh")[1],
                            lr = 1e-3) {
  keras_model_sequential() |>
    layer_masking(mask_value = 0, input_shape = c(time_steps, n_variables)) |>
    layer_lstm(units = 256, activation = "tanh", return_sequences = TRUE) |>
    layer_dropout(rate = 0.2) |>
    layer_lstm(units = 128, activation = "tanh", return_sequences = TRUE) |>
    layer_lstm(units = bottleneck, activation = "tanh", return_sequences = TRUE) |>
    layer_attention_sum() |>
    layer_repeat_vector(n = time_steps) |>
    layer_lstm(units = bottleneck, activation = "tanh", return_sequences = TRUE,
               recurrent_dropout = 0.0) |>
    layer_lstm(units = 128, activation = "tanh", return_sequences = TRUE) |>
    layer_dropout(rate = 0.2) |>
    layer_lstm(units = 256, activation = "tanh", return_sequences = TRUE) |>
    time_distributed(layer_dense(units = n_variables, activation = out_activation)) |>
    compile(
      optimizer = optimizer_adam(learning_rate = lr, clipnorm = 1.0),
      loss = "mse"
    )
}


model1 <- function(tr_tensor3d) {
  time_steps <- dim(tr_tensor3d)[2]
  n_variables <- dim(tr_tensor3d)[3]
  
  seq_ae_model <- keras_model_sequential()
  seq_ae_model <- seq_ae_model |> 
    layer_lstm(units = 128, activation = 'tanh', return_sequences = TRUE, input_shape = c(time_steps, n_variables)) |> 
    layer_dropout(rate = 0.2) |> 
    layer_lstm(units = 64, activation = 'tanh', return_sequences = FALSE) |> 
    layer_lstm(units = 8, activation = 'tanh', return_sequences = FALSE) |> 
    layer_repeat_vector(n = time_steps) |> 
    layer_lstm(units = 8, activation = 'tanh', return_sequences = TRUE) |> 
    layer_lstm(units = 64, activation = 'tanh', return_sequences = TRUE) |> 
    layer_dropout(rate = 0.2) |> 
    layer_lstm(units = 128, activation = 'tanh', return_sequences = TRUE) |> 
    time_distributed(layer_dense(units = n_variables, activation = "sigmoid")) |> 
    compile(optimizer = 'adam', loss = 'mse')
  
  return(seq_ae_model)
}

model2 <- function(tr_tensor3d) {
  time_steps <- dim(tr_tensor3d)[2]
  n_variables <- dim(tr_tensor3d)[3]
  
  seq_ae_model <- keras_model_sequential()
  seq_ae_model <- seq_ae_model |> 
    layer_lstm(units = 256, activation = 'tanh', return_sequences = TRUE, input_shape = c(time_steps, n_variables)) |> 
    layer_dropout(rate = 0.2) |> 
    layer_lstm(units = 64, activation = 'tanh', return_sequences = TRUE) |> 
    layer_lstm(units = 8, activation = 'tanh', return_sequences = FALSE) |> 
    layer_repeat_vector(n = time_steps) |> 
    layer_lstm(units = 8, activation = 'tanh', return_sequences = TRUE) |> 
    layer_lstm(units = 64, activation = 'tanh', return_sequences = TRUE) |> 
    layer_dropout(rate = 0.2) |> 
    layer_lstm(units = 256, activation = 'tanh', return_sequences = TRUE) |> 
    time_distributed(layer_dense(units = n_variables, activation = "sigmoid")) |> 
    compile(optimizer = 'adam', loss = 'mse')
  
  return(seq_ae_model)
}

model3 <- function(tr_tensor3d) {
  time_steps <- dim(tr_tensor3d)[2]
  n_variables <- dim(tr_tensor3d)[3]
  
  seq_ae_model <- keras_model_sequential()
  seq_ae_model <- seq_ae_model |> 
    layer_lstm(units = 16, activation = 'tanh', return_sequences = TRUE, input_shape = c(time_steps, n_variables)) |> 
    layer_dropout(rate = 0.2) |> 
    layer_lstm(units = 4, activation = 'tanh', return_sequences = FALSE) |> 
    layer_repeat_vector(n = time_steps) |> 
    layer_lstm(units = 4, activation = 'tanh', return_sequences = TRUE) |> 
    layer_dropout(rate = 0.2) |> 
    layer_lstm(units = 16, activation = 'tanh', return_sequences = TRUE) |> 
    time_distributed(layer_dense(units = n_variables, activation = "sigmoid")) |> 
    compile(optimizer = 'adam', loss = 'mse')
  
  return(seq_ae_model)
}

model4 <- function(tr_tensor3d) {
  time_steps <- dim(tr_tensor3d)[2]
  n_variables <- dim(tr_tensor3d)[3]
  
  seq_ae_model <- keras_model_sequential()
  seq_ae_model <- seq_ae_model |> 
    layer_lstm(units = 128, activation = 'tanh', return_sequences = TRUE, input_shape = c(time_steps, n_variables)) |> 
    layer_dropout(rate = 0.2) |> 
    layer_lstm(units = 32, activation = 'tanh', return_sequences = FALSE) |> 
    layer_lstm(units = 8, activation = 'tanh', return_sequences = FALSE) |> 
    layer_repeat_vector(n = time_steps) |> 
    layer_lstm(units = 8, activation = 'tanh', return_sequences = TRUE) |> 
    layer_lstm(units = 32, activation = 'tanh', return_sequences = TRUE) |> 
    layer_dropout(rate = 0.2) |> 
    layer_lstm(units = 128, activation = 'tanh', return_sequences = TRUE) |> 
    time_distributed(layer_dense(units = n_variables, activation = "sigmoid")) |> 
    compile(optimizer = 'adam', loss = 'mse')
  
  return(seq_ae_model)
}

model5 <- function(tr_tensor3d) {
  time_steps <- dim(tr_tensor3d)[2]
  n_variables <- dim(tr_tensor3d)[3]
  
  seq_ae_model <- keras_model_sequential()
  seq_ae_model <- seq_ae_model |> 
    layer_lstm(units = 128, activation = 'tanh', return_sequences = TRUE, input_shape = c(time_steps, n_variables)) |> 
    layer_dropout(rate = 0.2) |> 
    layer_lstm(units = 8, activation = 'tanh', return_sequences = FALSE) |> 
    layer_repeat_vector(n = time_steps) |> 
    layer_lstm(units = 8, activation = 'tanh', return_sequences = TRUE) |> 
    layer_dropout(rate = 0.2) |> 
    layer_lstm(units = 128, activation = 'tanh', return_sequences = TRUE) |> 
    time_distributed(layer_dense(units = n_variables, activation = "sigmoid")) |> 
    compile(optimizer = 'adam', loss = 'mse')
  
  return(seq_ae_model)
}

```



# 5. Training Data Preparation and Model Training

The training data, representing normal, healthy forest plots, gets reshaped into a 3D tensor before it is used to train the model. At first, the imputation methods were also incorporated into the seq_len loop. However, this turned out to be too difficult for the PC, so it was done individually.



**Kalman Run Model 1**
```{r}
# Training data
imputation_method <- fill_gaps_kal
train_interpolated <- kal_interpolated

for (seq_len in c(4, 8, 12, 24)) {
 
  # Nesting training data 
  nested <- train_interpolated |> group_by(coords) |> nest()
  
  # Converting it to one long time series 
  n <- length(nested$coords)
  f <- data.frame()
  
  for (i in 1:n) {
    d <- as.data.frame(nested$data[i])
    f <- rbind(f, d)
  }
  
  # Train dataframe 
  train_df <- f |> dplyr::select(all_of(names), -time)  # Removing time column

  # Scaler is trained on the training data 
  scaler <- preProcess(train_df, method = c("range"))

  # 3D tensor
  tr_df3d <- train_df |> as.data.frame()
  tr_cut_length <- floor(nrow(tr_df3d) / seq_len) * seq_len  # Length must be divisible by seq_len
  tr_df_length <- tr_df3d[1:tr_cut_length, ]
  tr_df_length <- predict(scaler, tr_df_length)
  tr_tsor_long3d <- long2tens3d_no_scaler(tr_df_length, tr_cut_length / seq_len)

  # Model Definition and Training
  remove(ae_seq_1)

  # Defining the model
  ae_seq_1 <- model1(tr_tensor3d = tr_tsor_long3d)

  # Fitting the model
  tens_training <- ae_seq_1 |> fit(tr_tsor_long3d, tr_tsor_long3d, epochs = 850, batch_size = 64, validation_split = 0.2, verbose = 0)
  path <- paste0(path_root, "Data/models/clean/ae1_kal_", seq_len, "_v1.h5")
  save_model_hdf5(ae_seq_1, path)
}

```






**MA durchlauf model 1**
```{r}
# Training data
imputation_method <- fill_gaps_ma
train_interpolated <- ma_interpolated

for (seq_len in c(4, 8, 12, 24)) {
 
  # Nesting training data 
  nested <- train_interpolated |> group_by(coords) |> nest()
  
  # Converting it to one long time series 
  n <- length(nested$coords)
  f <- data.frame()
  
  for (i in 1:n) {
    d <- as.data.frame(nested$data[i])
    f <- rbind(f, d)
  }
  
  # Train dataframe 
  train_df <- f |> dplyr::select(all_of(names), -time)  # Removing time column

  # Scaler is trained on the training data 
  scaler <- preProcess(train_df, method = c("range"))

  # 3D tensor
  tr_df3d <- train_df |> as.data.frame()
  tr_cut_length <- floor(nrow(tr_df3d) / seq_len) * seq_len  # Length must be divisible by seq_len
  tr_df_length <- tr_df3d[1:tr_cut_length, ]
  tr_df_length <- predict(scaler, tr_df_length)
  tr_tsor_long3d <- long2tens3d_no_scaler(tr_df_length, tr_cut_length / seq_len)

  # Model Definition and Training
  remove(ae_seq_1)

  # Defining the model
  ae_seq_1 <- model1(tr_tensor3d = tr_tsor_long3d)

  # Fitting the model
  tens_training <- ae_seq_1 |> fit(tr_tsor_long3d, tr_tsor_long3d, epochs = 850, batch_size = 64, validation_split = 0.2, verbose = 0)

  path <- paste0(path_root, "Data/models/clean/ae1_ma_", seq_len, "_v1.h5")
  save_model_hdf5(ae_seq_1, path)
}


```




**SGMA Run Model1** 
```{r}
# Training data
imputation_method <- fill_gaps_sg_ma
train_interpolated <- sg_ma_interpolated

for (seq_len in c(4, 8, 12, 24, 52)) {
 
  # Nesting training data 
  nested <- train_interpolated |> group_by(coords) |> nest()
  
  # Converting it to one long time series 
  n <- length(nested$coords)
  f <- data.frame()
  
  for (i in 1:n) {
    d <- as.data.frame(nested$data[i])
    f <- rbind(f, d)
  }
  
  # Train dataframe 
  train_df <- f |> dplyr::select(all_of(names), -time)  # Removing time column

  # Scaler is trained on the training data 
  scaler <- preProcess(train_df, method = c("range"))

  # 3D tensor
  tr_df3d <- train_df |> as.data.frame()
  tr_cut_length <- floor(nrow(tr_df3d) / seq_len) * seq_len  # Length must be divisible by seq_len
  tr_df_length <- tr_df3d[1:tr_cut_length, ]
  tr_df_length <- predict(scaler, tr_df_length)
  tr_tsor_long3d <- long2tens3d_no_scaler(tr_df_length, tr_cut_length / seq_len)

  # Model Definition and Training
  remove(ae_seq_1)

  # Defining the model
  ae_seq_1 <- model1(tr_tensor3d = tr_tsor_long3d)

  # Fitting the model
  tens_training <- ae_seq_1 |> fit(tr_tsor_long3d, tr_tsor_long3d, epochs = 850, batch_size = 64, validation_split = 0.2, verbose = 0)

  path <- paste0(path_root, "Data/models/clean/ae1_sgma_", seq_len, "_v1.h5")
  save_model_hdf5(ae_seq_1, path)
}

```


# 6 Model Architecture Selection 

To reduce repetitive code, a function is designed that:

**(1)** Creates a 3D tensor from the training data.

**(2)** Adjusts the selected model architecture to the training data shape.

**(3)** Fits the model on the training data.

The result is a fitted model


**Function: **
```{r}
# Training data
model_architecture_fun <- function(train_interpolated, seq_len, model, names, epochs=120) {
  
  # Nesting training data 
  nested <- train_interpolated |> group_by(coords) |> nest()
  
  # Converting it to one long time series 
  n <- length(nested$coords)
  f <- data.frame()
  
  for (i in 1:n) {
    d <- as.data.frame(nested$data[i])
    f <- rbind(f, d)
  }
  
  # Train dataframe 
  train_df <- f |> dplyr::select(all_of(names), -time)  # Removing time column

  # Scaler is trained on the training data 
  scaler <- preProcess(train_df, method = c("range"))

  # 3D tensor
  tr_df3d <- train_df |> as.data.frame()
  tr_cut_length <- floor(nrow(tr_df3d) / seq_len) * seq_len  # Length must be divisible by seq_len
  tr_df_length <- tr_df3d[1:tr_cut_length, ]
  tr_df_length <- predict(scaler, tr_df_length)
  tr_tsor_long3d <- long2tens3d_no_scaler(tr_df_length, tr_cut_length / seq_len)

  # Model Definition and Training
  remove(ae_seq_1)  # Removing the model variable in case it is already defined as a different model
  
  # Defining the model
  ae_seq_1 <- model(tr_tensor3d = tr_tsor_long3d)
  
  # Fitting the model
  tens_training <- ae_seq_1 |> fit(tr_tsor_long3d, tr_tsor_long3d, epochs = epochs, batch_size = 64, validation_split = 0.2, verbose = 1)

  return(ae_seq_1)
}

```



### Applying the function on the different models for: 
 
**SGMA -- seq_len = 4**
**SGMA -- seq_len = 24**

```{r}

ae2_4_v1<-model_architecture_fun(train_interpolated, 4, model2, vars1)
path = paste0(path_root, "Data/models/clean/ae2_4_v1.h5")
save_model_hdf5(ae2_4_v1, path)

ae3_4_v1<-model_architecture_fun(train_interpolated, 4, model3, vars1)
path = paste0(path_root, "Data/models/clean/ae3_4_v1.h5")
save_model_hdf5(ae3_4_v1, path)

ae4_4_v1<-model_architecture_fun(train_interpolated, 4, model4, vars1)
path = paste0(path_root, "Data/models/clean/ae4_4_v1.h5")
save_model_hdf5(ae4_4_v1, path)

ae5_4_v1<-model_architecture_fun(train_interpolated, 4, model5, vars1)
path = paste0(path_root, "Data/models/clean/ae5_4_v1.h5")
save_model_hdf5(ae5_4_v1, path)


  
#############################
#         window size 24



ae2_24_v1<-model_architecture_fun(train_interpolated, 24, model2, names = vars1)
path = paste0(path_root, "Data/models/clean/ae2_24_v1.h5")
save_model_hdf5(ae2_24_v1, path)

ae3_24_v1<-model_architecture_fun(train_interpolated, 24, model3, names = vars1)
path = paste0(path_root, "Data/models/clean/ae3_24_v1.h5")
save_model_hdf5(ae3_24_v1, path)


ae4_24_v1<-model_architecture_fun(train_interpolated, 24, model4, names = vars1)
path = paste0(path_root, "Data/models/clean/ae4_24_v1.h5")
save_model_hdf5(ae4_24_v1, path)

ae5_24_v1<-model_architecture_fun(train_interpolated, 24, model5, names = vars1)
path = paste0(path_root, "Data/models/clean/ae5_24_v1.h5")
save_model_hdf5(ae5_24_v1, path)

```



# 7 Model Variable input Selection:  


**SG_MA is the best method and is used for all variables**

for window 24: model3, model2

**(16/8)**
**(256/64/8)**


for window 4: model2

**(256/64/8)**

 **Model2**
 
 
 
 
 

# sg_ma durchlauf model 2
```{r}
# Training data
train_interpolated <- sg_ma_interpolated

# Variable lists and names
var_list <- list(vars2, vars3, vars4, vars5)
var_names <- list("v2", "v3", "v4", "v5")

# AE2, window size = 4
for (var in 1:4) {
  variables <- var_list[[var]]
  mod <- model_architecture_fun(train_interpolated, 4, model2, names = variables)
  path <- paste0(path_root, "Data/models/clean/other Vars/ae2_w4_sg_ma_", var_names[[var]], ".h5")
  save_model_hdf5(mod, path)
}

# AE2, window size = 24
for (var in 1:4) {
  variables <- var_list[[var]]
  mod <- model_architecture_fun(train_interpolated, 24, model2, names = variables)
  path <- paste0(path_root, "Data/models/clean/other Vars/ae2_w24_sg_ma_", var_names[[var]], ".h5")
  save_model_hdf5(mod, path)
}

# AE3, window size = 24
for (var in 1:4) {
  variables <- var_list[[var]]
  mod <- model_architecture_fun(train_interpolated, 24, model3, names = variables)
  path <- paste0(path_root, "Data/models/clean/other Vars/ae3_w24_sg_ma_", var_names[[var]], ".h5")
  save_model_hdf5(mod, path)
}


```




```{r}
#var 5 overwrite

var<-4
  variables <- var_list[[var]]
  mod <- model_architecture_fun(train_interpolated, 4, model2, names = variables)
  
  xxxy<-3
  path <- paste0(path_root, "Data/models/clean/other Vars/ae2_w4_sg_ma_", var_names[[var]], "ovr.h5")
  save_model_hdf5(mod, path)


# AE2, window size = 24
  variables <- var_list[[var]]
  mod <- model_architecture_fun(train_interpolated, 24, model2, names = variables)
  path <- paste0(path_root, "Data/models/clean/other Vars/ae2_w24_sg_ma_", var_names[[var]], "ovr.h5")
  save_model_hdf5(mod, path)

# AE3, window size = 24
  variables <- var_list[[var]]
  mod <- model_architecture_fun(train_interpolated, 24, model3, names = variables)
  path <- paste0(path_root, "Data/models/clean/other Vars/ae3_w24_sg_ma_", var_names[[var]], "ovr.h5")
  save_model_hdf5(mod, path)




```


```{r}
attention_model <- function(time_steps, n_variables, D = 64, lr = 1e-3,
                            out_activation = c("sigmoid","linear","tanh")[1]) {
  inputs <- layer_input(shape = c(time_steps, n_variables))

  # Encoder
  enc_seq <- inputs %>%
    layer_masking(mask_value = 0) %>%
    layer_lstm(256, return_sequences = TRUE, activation = "tanh") %>%
    layer_dropout(0.2) %>%
    layer_lstm(128, return_sequences = TRUE, activation = "tanh") %>%
    layer_lstm(D,   return_sequences = TRUE, activation = "tanh")   # [B,T,D]

  # Attention pooling to a 1-step context (bottleneck)
  q <- enc_seq %>%
    layer_global_average_pooling_1d() %>%   # [B,D]
    layer_dense(units = D, activation = "tanh") %>%
    layer_reshape(target_shape = c(1, D))   # [B,1,D]

  context <- layer_attention(list(q, enc_seq), use_scale = TRUE)  # [B,1,D]
  z <- context %>% layer_flatten()                                # [B,D]

  # Decoder
  y <- z %>%
    layer_repeat_vector(time_steps) %>%
    layer_lstm(D,   return_sequences = TRUE, activation = "tanh") %>%
    layer_lstm(128, return_sequences = TRUE, activation = "tanh") %>%
    layer_dropout(0.2) %>%
    layer_lstm(256, return_sequences = TRUE, activation = "tanh") %>%
    time_distributed(layer_dense(units = n_variables, activation = out_activation))

  model <- keras_model(inputs, y)
  model %>% compile(optimizer = optimizer_adam(learning_rate = lr, clipnorm = 1.0),
                    loss = "mse")
  model
}

```



```{r}



names<-c("time","b2","b3","b4","b5","b6","b7","b8","b8a","b11","b12")
names_nt<-c("b2","b3","b4","b5","b6","b7","b8","b8a","b11","b12")
train_f<-train|> filter(coords %in% unique(train$coords)[1:1000])
train_interpolated<-gf_impute_df_sgma2_schnell(train_f,group_key = "coords",order = 3,aggregate = "week",padding = F,mask=0, k_ma = 8,dt = FALSE)

scaler <- preProcess(train_interpolated[names_nt], method = c("range"))
train_interpolated_scaled<-train_interpolated
train_interpolated_scaled[names_nt] <-  predict(scaler, train_interpolated[names_nt]) 

train_tens<-train_interpolated_scaled|>select(-dt)|>create_dataset_stepwise_opt(seq_len = 52,step = 30)

# train_tens: torch_tensor [batch, time, features]
x_arr <- as.array(train_tens$to(dtype = torch_float(), device = "cpu"))

# (Optional) replace NaN if present; Keras can’t handle NaN
x_arr[!is.finite(x_arr)] <- 0

# ensure float32 for TF
#storage.mode(x_arr) <- "single"
#x_arr[1,,]

att_model <- attention_model(time_steps = 52, n_variables = dim(x_arr)[3], D = 64, out_activation = "linear")

history <- att_model %>% fit(
  x = x_arr,
  y = x_arr,                 # AE target = input
  batch_size = 64,
  epochs = 50,
  validation_split = 0.1,
  callbacks = list(
    callback_early_stopping(monitor = "val_loss", patience = 5, restore_best_weights = TRUE)
  )
)
# 
# 
# attention_model<-model_architecture_fun(train_tens, 52, attention_model, names = c("coords", names))

path = "C:/Users/maxki/Documents/Paper/new_models/models/keras_attention.h5"

save_model_hdf5(att_model, path)
# keras::save_model_weights_tf(att_model, filepath = 'C:/Users/maxki/Documents/Paper/new_models/models/keras_attention.keras')
```


# Infernez
```{r}
#test1<-test1|>rename(coords= tree_id)

test_imp<-gf_impute_df_sgma2_schnell(test1,group_key = "coords",order = 3,aggregate = "week",padding = F,mask=NA, dt = F, k_ma = 8)


test_imp_scaled<-test_imp
test_imp_scaled[names_nt] <-  predict(scaler, test_imp_scaled[names_nt]) 



test_tens<-test_imp_scaled|>create_dataset_stepwise_opt(seq_len = 52,step = 52)
test_tens_arr <- as.array(test_tens$to(dtype = torch_float(), device = "cpu"))
test_tens_arr

test_tens_arr_out<-predict(att_model, test_tens_arr)
test_tens_arr_out



test_tens

test_tens_arr_out[1,,]



```


```{r}
ref_idx <- test_imp_scaled %>%
  arrange(coords, time) %>%
  group_by(coords) %>%
  mutate(row_id = row_number(),
         n_full = n() %/% 52,
         keep_n = n_full * 52) %>%
  filter(row_id <= keep_n) %>%
  ungroup() %>%
  select(coords, time)

# run model
y_hat <- predict(att_model, test_tens_arr)  # [W, 52, F]

# flatten windows back to rows in same order as ref_idx
W <- dim(y_hat)[1]; Tt <- dim(y_hat)[2]; Ff <- dim(y_hat)[3]
pred_mat <- do.call(rbind, lapply(seq_len(W), function(i) y_hat[i,,]))  # [W*52, F]
pred_df  <- as.data.frame(pred_mat)
colnames(pred_df) <- names_nt  # your variable names

pred_with_meta <- bind_cols(ref_idx, pred_df) %>%
  arrange(coords, time)
```


```{r}

# new_names<- lapply(colnames(pred_with_meta)[3:12], function(x) paste0("pred_",x)) 
# colnames(pred_with_meta)[3:12]<-new_names

alldata<-left_join(test_imp_scaled, pred_with_meta, by = c("coords", "time"))

X <- as.matrix(alldata[, 3:12])
Y <- as.matrix(alldata[, 13:22])

stopifnot(ncol(X) == ncol(Y))   # safety
alldata$rec_err <- rowMeans(abs(X - Y), na.rm = TRUE)



alldata|>
  filter(coords==unique(alldata$coords)[200])|>
  ggplot(aes(x=as_date(time)))+
  geom_line(aes(y= b12))+
  geom_line(aes(y= pred_b12), col= "orange")+
  geom_line(aes(y= rec_err), col= "red")+
  theme_bw()


```




```{r}

reconstruct_function<-function(MODEL,scaler,test_data, imputation_method ,window_size){
  # the actual function for individual time series 
  rec_function<-function(x){
    x <- x |> as.data.frame() |> 
      select(all_of(names))
    x <- imputation_method(x)  # Apply the imputation method here
    
    # 3d tensor 
    cut_length <- floor(nrow(x)/window_size)*window_size # because the length must be devidable by the window_size
    test_df_length<-x[1:cut_length,]
    # test_df_length will be the reference df. It is scaled here
    ref_df<- predict(scaler, test_df_length)
    
    test_df_length_nt<-test_df_length|> dplyr::select(-time)
    test_df_length_nt<-predict(scaler, test_df_length_nt)
    test_tens3d<-long2tens3d_no_scaler(test_df_length_nt, (cut_length/window_size))
    
    # predict
    reconstructed_sequences <-  predict(MODEL, test_tens3d, verbose = 1)
    
    
    
    # combining reconstrction and scaled original data in a dataset
    vector_list <- list()
    for(i in seq(2:length(names))){
      j<-i+1
      x<-c(t(reconstructed_sequences[,,i]))
      vector_list[[paste0("rec_", names[j])]] <- x
    }
    rec_df <- as.data.frame(vector_list)
    Lossma <- rowMeans(abs(rec_df - ref_df[2:ncol(ref_df)])) 
    total_df<-cbind(ref_df, rec_df, rec_err = Lossma)
    
    return(total_df)
    
  }
 
  ## Applying the function
  
  
  nest<-test_data|> group_by(coords)|> nest()
  
  reconstructed_dataframe <- nest|> 
    mutate(data= map(data, rec_function))|>
    unnest()|> ungroup()|> as.data.frame()
  
  
# if there is a space in the coords column:   
  reconstructed_dataframe <- reconstructed_dataframe |> 
    mutate(coords = ifelse(grepl("/\\s", coords), gsub("/\\s+", "/", coords), coords))
  
  
  
  print("Done")
  return(reconstructed_dataframe)
}


```


